\Universita{University of Verona}
\Dipartimento{Department of Informatics}
\CorsoDiLaurea{Master's Degree in Data Science}
\AnnoAccademico{Academic Year 2021--2022}
\Titolo{Leveraging LLMs with RAG to recommend Points of Interest to tourists}
\Relatore{Prof. Niccolò \textsc{Marastoni}}
\RelatoreLabel{Supervisor}
\CandidatoLabel{Graduate Student}
\Candidato{Patrick \textsc{Hamzaj}} 
\Matricola{VR474246}

\begin{titlepage}
    \newgeometry{left=3cm, right=3cm, bottom=2cm, top =3cm} 
    \pagestyle{empty}
    \makefrontpage
    \restoregeometry
\end{titlepage}

\newpage
\thispagestyle{empty}
\mbox{}  % This ensures an empty page
\newpage


\frontmatter
% dedica
\null\vspace{\stretch{1}}
\begin{flushright}
    \textit{Ad Artur, Alketa, Cristal, Melissa, a tutta la mia famiglia, e tutti i miei amici.}

\end{flushright}
\vspace{\stretch{2}}\null

%ringraziamenti
\chapter*{Acknowledgments}
First and foremost, I would like to express my deepest gratitude to Professor Niccolò Marastoni, whose unwavering support throughout my entire academic journey has been invaluable--not only during the writing of this thesis but also in all aspects of my studies.

I am also forever grateful to my parents Artur and Alketa, who have always been there for me no matter what, providing constant encouragement and unconditional love. My sister Cristal, for being such a lovely sister, I owe you the world.

A special thanks goes to my girlfriend Melissa, with whom I have shared countless experiences. Her support has meant the world to me--even though she sometimes playfully protests when I "borrow" her hairdryer adapter.

I want to thank Federico, affectionately known as Chicco, with which I shared this academic journey and an apartment.

A special thanks to La Saletta, I grew up with all of you, sharing a special place and experiences, I always have so much fun with you. My friends, I am forever grateful for having you.

\begin{abstract}
This thesis explores the practical integration of Large Language Models (LLMs) with innovative strategies that harness their generative capabilities in limited resource environments. The study focuses on the use of the open-source LLaMA 3.1 8B-Instruct model, enhanced through strategies such as prompt engineering, retrieval-augmented generation (RAG), and Agent AI to propose an application that leverages data from datasets involving check-ins of tourists to Verona's attractions and weather forecasts to provide contextually relevant recommendations. The project, through a comprehensive evaluation involving conversational analysis, assesses system performance, identifies limitations, and suggests directions for future improvements.
\end{abstract}

\tableofcontents

\printnomenclature
\addcontentsline{toc}{chapter}{Nomenclature}

\mainmatter
\chapter{Introduction}
\markboth{\MakeUppercase{Introduction}}{}
\input{capitoli/intro}